State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-10 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
some edition for plots caption

   	\caption{Kernel centred at zero as a function of space.} 

@@

\centering
\begin{tabular}{c|c|c}
	\hline\hline Symbol & Value & Units \\
	\hline\hline
	$\zeta$ & 0.01 & ms$^{-1}$\\
	$f_{max}$ & 10 & spikes/s \\
	$\varsigma$ & 0.8 & spike/mV\\
	$v_0$ & 2 & mV\\
	$\boldsymbol{\theta}$ & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\right]^{\top}$ & synaptic strength\\
	$\sigma_{\psi,1}$ & 1.8 & inner connectivity kernel width\\
	$\sigma_{\psi,2}$ & 2.4 & middle connectivity kernel width\\
	$\sigma_{\psi,3}$ & 6 & outer connectivity kernel width\\
	$\Sigma$ & & field disturbance covariance \\
	sensor parameters & & \\
	$\Delta$ & 0.5 & space step \\
	$T_s$ & 1 & ms time step \\ 
	\hline\hline
\end{tabular}\label{tab:Model Parameters}
\caption{Parameters for neural field model used to generate data to demonstrate the estimation framework.}
\end{table*}
The sensors were spaced in a regular square \parham{14 $\times$ 14} grid, the sensor width $\sigma^2_m$ was \parham{0.81} which equals \parham{$1.5$}~mm full width at half maximum, with spacing $\Delta_y$ equal to \parham{$1.5$}~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design that is sufficient to guarantee an accurate representation of the field can be achieved using the frequency analysis from Section~\ref{SpectralAnalysisSection} given access to the true field. 
\begin{figure}[th]\label{fig:experimental design}
   		\subfigure[][]{\includegraphics{./Graph/ExperimentFigureEEGpy.pdf}} \\
   		\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePSD.pdf}} \\
   	\caption{Example of experimental design and model output.  \textbf{a} Example of data generated by the model through a subset of the observations. \textbf{b}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
% \begin{figure}[th]\label{fig:experimental design}
% \begin{center}
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePSDPy.pdf}}
% \end{center}
% \caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
% \end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameters estimates over 100
realisations. \textbf{a.} The distribution of the synaptic time constant estimates (true parameter: $\zeta=100$). \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameters estimates (true parameter: $\theta_0 = 10$). \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates (true parameter: $\theta_1 = -8$). \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates (true parameter $\theta_3 = 0.5$).}
\label{fig:Parameters}
\begin{figure}[th]
\includegraphics[scale=1]{./Graph/convergence.pdf}
\caption{Convergence of the parameters averaged over 100 runs. }
\label{fig:ParametersConvergence}
\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT}(a) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \parham{0.24} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \parham{2.05}mm. This confirms our separation of \parham{1.5}mm was enough to prevent problems associated with aliasing.

@@

\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain. \dean{Note: needs to be changed to absolute values if we include it.} \dean{Note: Power and dB needs to be non-italicised.}}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20).}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics[scale=1]{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics[scale=1]{./Graph/FFTEstimatedField.pdf}}\\
\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field resulting from the basis function decomposition. As a result of the band limiting properties of the sensors and basis functions the reconstructed shows a narrower bandwidth.}
\label{fig:FFTTrueEstimate}
\end{figure}

  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region). \dean{David suggested for us to use percentage error, and use a hard threshold of points close to zero to stop problems with divide by zeros. Also, it might be good to show the data points along the line, so we can see the number of iterations it takes to stabilise.}} 
\label{fig:RMSE}
   \end{figure}
\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[scale=1]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[scale=1]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[scale=1]{./Graph/EstimationError90.pdf}}\\
\caption{Examples of the neural fields. \textbf{a.} The true field from the non-reduced model the was used to generate data. \textbf{b.} The reconstructed estimated field of the reduced model. \textbf{c.} The absolute error between the true and reconstructed fields.}
\label{fig:FieldEstimate}
\end{figure*}

>>>>

   	\caption{A mexican-hat connectivity kernel, $w(\mathbf{r},\mathbf{r'})$ used in this paper. The function $w(.)$ is rotationally symetric (isotropic) about zero, hence a cross section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.} 

@@

\centering
\begin{tabular}{c|c|c}
	\hline\hline Symbol & Value & Units \\
	\hline\hline
	$\zeta$ & 0.01 & ms$^{-1}$\\
	$f_{max}$ & 10 & spikes/s \\
	$\varsigma$ & 0.8 & spike/mV\\
	$v_0$ & 2 & mV\\
	$\boldsymbol{\theta}$ & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & synaptic strength\\
	$\sigma_{\psi,1}$ & 1.8 & inner connectivity kernel width\\
	$\sigma_{\psi,2}$ & 2.4 & middle connectivity kernel width\\
	$\sigma_{\psi,3}$ & 6 & outer connectivity kernel width\\
	$\Sigma$ & & field disturbance covariance \\
	sensor parameters & & \\
	$\Delta$ & 0.5 & space step \\
	$T_s$ & 1 & ms time step \\ 
	\hline\hline
\end{tabular}\label{tab:Model Parameters}
\caption{Parameters for neural field model used to generate data to demonstrate the estimation framework.}
\end{table*}
The sensors were spaced in a regular square \parham{14 $\times$ 14} grid, the sensor width $\sigma^2_m$ was \parham{0.81} which equals \parham{$1.5$}~mm full width at half maximum, with spacing $\Delta_y$ equal to \parham{$1.5$}~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design that is sufficient to guarantee an accurate representation of the field can be achieved using the frequency analysis from Section~\ref{SpectralAnalysisSection} given access to the true field. 
\begin{figure}[th]\label{fig:experimental design}
   	\begin{center}
    		\subfigure[][]{\includegraphics{./Graph/ExperimentFigureEEGPy.pdf}} \\
   		\subfigure[][]{\includegraphics{./Graph/ExperimentFigurePSDPy.pdf}} \\
   	\end{center}
   	\caption{Example of experimental design and model output.  \textbf{a} Example of data generated by the non-reduced model through a subset of the observations. The sensors were spaced in a regular square $14 \times 14$ grid. \textbf{b}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
   \end{figure}

\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameters estimates over 100
realisations. \textbf{a.} The distribution of $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the synaptic time constant (true parameter: $\xi=0.9$). \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameters estimates (true parameter: $\theta_0 = 10$). \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates (true parameter: $\theta_1 = -8$). \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates (true parameter $\theta_3 = 0.5$). The dotted line shows the location of the true parameter in each case.}
\label{fig:Parameters}
\end{figure}
\begin{figure}[th]
\includegraphics[scale=1]{./Graph/convergence.pdf}
\caption{The absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations. The errors fall below $10^{-4}$ after six iterations. }
\label{fig:ParametersConvergence}
\end{figure}
\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT}(a) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \parham{0.24} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \parham{2.05}mm. This confirms our separation of \parham{1.5}mm was enough to prevent problems associated with aliasing.

@@

\begin{figure*}
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity kernel estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain. \dean{Note: needs to be changed to absolute values if we include it.} \dean{Note: Power and dB needs to be non-italicised.}}
\label{fig:KernelEstimates}
\end{figure*}
\begin{figure}[th]
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). The dotted line shows the location of the true ratio in each case.}
\end{figure}
\label{fig:ParametersRatio}
\begin{figure}[th]
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics[scale=1]{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics[scale=1]{./Graph/FFTEstimatedField.pdf}}\\
\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field resulting from the basis function decomposition. As a result of the band limiting properties of the sensors and basis functions the reconstructed field shows a narrower bandwidth.}
\label{fig:FFTTrueEstimate}
\end{figure}

  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region). The error remains consitent over time. \dean{David suggested for us to use percentage error, and use a hard threshold of points close to zero to stop problems with divide by zeros. Also, it might be good to show the data points along the line, so we can see the number of iterations it takes to stabilise.}} 
\label{fig:RMSE}
   \end{figure}
\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[scale=1]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[scale=1]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[scale=1]{./Graph/EstimationError90.pdf}}\\
\caption{Examples of the neural fields. \textbf{a.} The true field from the non-reduced model which was used to generate data. \textbf{b.} The reconstructed estimated field of the reduced model, showing the effect of the basis function decomposition on the frequency content of the estimated field. \textbf{c.} The absolute error between the true and reconstructed fields, the error shows high frequency behaviour.}
\label{fig:FieldEstimate}
\end{figure*}
